#Reproducible Research Course Project NO.1 

Loading and preprocessing the data 
```{r}
setwd("/Users/LucasXin/Documents/R/Data Science Coursera/Reproducible-Research")
activity<-read.csv("activity.csv",header=TRUE,sep=",",na.strings="NA",colClasses=c(date="Date"))
str(activity)
head(activity)
```

It is confirmed that there are 3 variables and 17,568 observations, which was expected. Now that the data hs been loaded, we can do the analysis

## What is the mean total number of steps taken per day? 
First, the total number of steps taken perday is calculated using the dplyr library
```{r}
library(dplyr)
total_step_day<-activity %>% 
                group_by(date) %>%
                summarize(total.steps=sum(steps, na.rm=FALSE))
head(total_step_day)
```

Below is the distribution of the total number of stpes taken per day (missing values ignored)
```{r}
hist(total_step_day$total.steps,
     main="Distribution of steps by Day",
     xlab="Total Nubmer of Steps per Day",
     ylab="Frequency",
     breaks=20, col="red")
```
The mean total number of steps taken per day can be calcualted as: 
```{r}
mean(total_step_day$total.steps,na.rm=TRUE)
```
and the median 
```{r}
median(total_step_day$total.steps,na.rm=TRUE)
```

## What is the average daily activity pattern? 
To answer this question, the average number of steps per 5-min internval is compter 
```{r}
AVG_step<-activity %>%
        group_by(interval) %>%
        summarize(avg.steps=mean(steps,na.rm=TRUE))
head(AVG_step)
```
The figure below shows the average daily activity patteron.
```{r}
plot(strptime(sprintf("%04d", AVG_step$interval), format="%H%M"),
     AVG_step$avg.steps, type = "l", 
     main = "Average Daily Activity", 
     xlab = "Time of Day (HH:MM)", 
     ylab = "Average Number of Steps")
```
## Imputing missing values
### Summarize all the missing values
```{r} 
sum(is.na(activity$steps))
```
### Fill missing values with NA
```{r}
data_full <- activity
nas <- is.na(data_full$steps)
avg_interval <- tapply(data_full$steps, data_full$interval, mean, na.rm=TRUE, simplify=TRUE)
data_full$steps[nas] <- avg_interval[as.character(data_full$interval[nas])]
```
### Check to see if the missing values are patched
```{r}
sum(is.na(data_full$steps))
```
### Calculate the number of steps 
```{r}
steps_full <- data_full %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print

library(ggplot2)
ggplot(steps_full, aes(x = steps)) +
  geom_histogram(fill = "red", binwidth = 1000) +
  labs(title = "Histogram of Steps per day, including missing values", x = "Steps per day", y = "Frequency")
```

## Are there differences in activity patterns between weekdays and weekends?

### create a new column with mutate from dplyr,and apply whether the day is weekend or weekday
```{r}
data_full <- mutate(data_full, weektype = ifelse(weekdays(data_full$date) == "Saturday" | weekdays(data_full$date) == "Sunday", "weekend", "weekday"))
data_full$weektype <- as.factor(data_full$weektype)
head(data_full)
```
Calculate the average steps in the 5-minute interval and use ggplot for making the time series of the 5-minute interval for weekday and weekend, and compare the average steps
```{r}
interval_full <- data_full %>%
  group_by(interval, weektype) %>%
  summarise(steps = mean(steps))
s <- ggplot(interval_full, aes(x=interval, y=steps, color = weektype)) +
  geom_line() +
  facet_wrap(~weektype, ncol = 1, nrow=2)
print(s)
```